---
title: "Review template"
author: "Yann"
date: "`r format(Sys.time(), '%d/%m/%y')`"
output:
  html_document:
    code_folding: hide

---
# Data cleaning checks {.tabset}
version 1.1.0.  
<br />

This review is separated into two parts *Data checks* and *Cleaning log*:  
- The first part will performed different check on the dataset, depending on the information available 
(time, outliers, shortest path, etc.).  
- The second part will look at how the cleaning was performed. It will first look if the values from the
cleaning log are correctly changed in the clean data, and then it will look at differences between the 
raw data and clean data and see if anything was not logged into the cleaning log.  
<br />

The summary can be found in the output folder.


```{r, message = F, warning = F}
library(tidyverse)
library(knitr)
library(kableExtra)
#install.packages("openxlsx")
library(openxlsx)
#install.packages("remotes")
#remotes::install_github("ellieallien/cleaninginspectoR")
source("utils/rendering.R")
source("utils/check_cleaning.R")
file.remove("outputs/dataset_with_check.xlsx", "outputs/review_cleaning_part2.xlsx", "outputs/review_cleaning_log_part1.xlsx")
```


```{r, message = F, warning = F}
library(readxl)
#The only section that should be changed.
file <- "C:/Users/lisa.reiners/OneDrive - ACTED/Documents/R-training/Codigos de R/review_ajustado_lisa/inputs/IRQ2301_Winter_cash_assistance_dataset.xlsx"
raww <- read_excel(file, guess = 10000, sheet = "Raw Data")

cleann <- read_excel(file, guess = 50000, sheet="Clean Data")
cleann$start<-strftime(cleann$start,"%Y-%m-%dT%H:%M:%S.000%z")
cleann$end<-strftime(cleann$end,"%Y-%m-%dT%H:%M:%S.000%z")


logg <- read_excel(file, guess = 50000,sheet = "Cleaning Log")
colnames(logg) <- logg[1,]
logg <- logg[-1, ]


dell <- read_excel(file, guess = 50000, sheet = "Deletions")

questions <- read_excel(file, sheet="Kobo")

choices <- read_excel(file, sheet = "Choice")

# Create a list to store datasets
datasetList <- list(raww = raww, cleann = cleann,logg = logg, dell = dell)

# Iterate through the datasets in the list
for (name_datasets in names(datasetList)) {
  dataset <- datasetList[[name_datasets]]  # Get the current dataset
  
  have_uuid <- FALSE  # Flag to track if there are duplicate uuid columns
  column_names <- colnames(dataset)  # Get all column names
  duplicate_columns <- duplicated(column_names)  # Find duplicate column names
  
  # Check if there are any duplicate columns
  if (any(duplicate_columns)) {
    duplicate_column_names <- column_names[duplicate_columns]  # Save the names of the duplicated columns
    
    # Iterate through the duplicated column names
    for (value in duplicate_column_names) {
      if (value == "uuid" || value == "_uuid" || value == "__uuid") {
        have_uuid <- TRUE
        print(paste("There are duplicated uuid columns. Please choose the correct one and delete the incorrect one. It is in the:", name_datasets))
      }
    }
  }
  
  # If there are no duplicate uuid columns
  if (!have_uuid) {
    column_found <- FALSE  # Flag to track if the column is found
    uuid_name_datasets <- NULL  # Variable to store the name of the UUID column
    
    # Iterate through the column names
    for (column in column_names) {
      if (column == "uuid" || column == "_uuid" || column == "__uuid") {
        uuid_name_datasets <- column
        column_found <- TRUE
        break  # Stop the loop if the column is found
      }
    }
    
    if (column_found) {
      #print(paste("The column 'uuid' is found in:", name_datasets, "dataset"))
      #print(uuid_name_datasets)
      
      # Create a dynamic variable name to store the uuid_name_datasets value
      assign(paste("uuid_", name_datasets, sep = ""), uuid_name_datasets)
      
    } else {
      print(paste("Column 'uuid' not found!", name_datasets, "dataset"))
    }
  }
}

#Prueba
#uuid_cleann
#uuid_logg

#names from the cleaning log
colname_logg <- colnames(logg)
var_logg <- NULL
old_value_logg <- NULL

for (col_name in colname_logg) {
  if (grepl("question", tolower(col_name))) {
    var_logg <- col_name
    #print(var_logg)
  }
  if (grepl("old", tolower(col_name))) {
    old_value_logg <- col_name
    #print(old_value_logg)
  }
  if (grepl("new", tolower(col_name))) {
    new_value_logg <- col_name
    #print(new_value_logg)
  }
}

#prueba
#new_value_logg

#####################################  
# uuid_raww <- "_uuid"
# uuid_cleann <- "_uuid"
# uuid_logg <- "uuid"
# uuid_dell <- "_uuid"
# 
# #names from the cleaning log
# var_logg <- "Question"
# old_value_logg <- "Old Value"
# new_value_logg <- "New Value"

#tests to include  TRUE / FALSE
check_for_duplicates <- TRUE
check_for_pii <- TRUE
check_for_time <- TRUE
check_for_shortest_path <- TRUE 
check_for_outliers <- TRUE 
check_for_logical <- TRUE
check_for_others <- TRUE
check_for_deletions <- TRUE
check_for_etc <- TRUE
check_for_cleaning_log <- TRUE
check_for_raw_clean_data <- TRUE
```

```{r, message = F, warning = F}
#create a binding key for the cleaning log
logg <- logg %>%
  mutate(binding = paste0(!!sym(uuid_logg), "-/-", !!sym(var_logg)))
list_log <- list()
```

## 1. Data checks {.tabset}

### Checks for duplicates 

```{r} 
if (check_for_duplicates) {
  duplicate_log <- cleaninginspectoR::find_duplicates(cleann, uuid_cleann) %>% 
    mutate(uuid = cleann[[uuid_cleann]][index]) 
  
  list_log$duplicate_log <- duplicate_log
  
  print_log(duplicate_log, "No duplicates found") 

}
```


### Checks for PII

```{r, warning = F, message = F}
if (check_for_pii) {
  pii_log <- cleaninginspectoR::sensitive_columns(cleann)

  list_log$pii_log <- pii_log
  
  print_log(pii_log, "No sensitive columns found") ### how this understand that pii is empty?
}
```
*NOTE*:   
- Only looks for the words "GPS, gps, phone, Latitude, Longitude, Phone" in the names of the 
dataset.  
- It does not check the value in those columns


### Check for time
```{r, warning = F}
lower_treshold <- 15
higher_threshold <- 100

if (check_for_time) {
  time_log <- cleaninginspectoR::check_time(cleann, duration_threshold_lower = lower_treshold,
                                          duration_threshold_upper = higher_threshold)
  
  list_log$time_log <- time_log
  
  print_log(time_log, "No time sensitive interviews found") 

}
```
<br />
**Note**:  
- Check time for lower threshold as `r lower_treshold` minutes and higher threshold as `r higher_threshold` minutes.


### Check for shortest path

```{r}
# take only select and integer to look at NA (removing text, dummies, notes, etc.)
if (check_for_shortest_path) {
  if (exists("questions")) {
    names_for_blanks_check <- questions %>%
      filter(grepl("select", type) | type == "integer") %>%
      pull(name)
    
    cleann$blanks <- cleann %>% 
      select(any_of(names_for_blanks_check)) %>% 
      check_blanks()
    
  } else {
    cleann$blanks <- cleann %>% 
      check_blanks()
  }
  
  cleann$blanks %>% boxplot()
}


cleann <- cleann[order(cleann$blanks, decreasing = TRUE), ]
```
  

**NOTE**:  
- This plot shows how many NA the interviews have. You should check values with lots of blanks. As
there is not a specific threshold, but you should check interviews for high percentage of blanks. 
*Dataset with the **blank** column is in the output folder.*

### Check for outliers

```{r}
if (check_for_outliers) {
  outlier_log <- cleaninginspectoR::find_outliers(cleann) %>% 
    mutate(uuid = cleann[[uuid_cleann]][index]) 

  list_log$outlier_log <- outlier_log
  
  print_log(outlier_log, "No outlier found") 

}
```




### Logical check 

```{r, warning = F, message = F}
# if (check_for_logical) {
# #to be done manually
# #######################################################################################################
# #if you have a table with 2 columns : "name" and "check", where "check" is the syntax of the r check #
# #you can try to un-comment this part #
# #######################################################################################################
# 
# logical_list_check <- readxl::read_excel("inputs/check_list.xlsx")
# 
# 
# 
# make_check <- function(xx) {
# message(xx[["name"]])
# cleann %>%
# mutate(!!sym(xx[["name"]]) := eval(parse(text = xx[["check"]]))) %>%
# pull(!!sym(xx[["name"]]))
# }
# 
# 
# 
# logical_check_table <- logical_list_check %>%
# apply(1, make_check) %>%
# as.data.frame() %>%
# `names<-`(logical_list_check$name)
# 
# 
# 
# cleann <- cbind(cleann, logical_check_table)
# 
# 
# 
# summary_logical_check <- logical_check_table %>%
# summarise(across(.fns = sum, na.rm=T)) %>%
# t() %>%
# cbind(logical_list_check,.) %>%
# `names<-`(c("name", "check", "n")) %>%
# arrange(desc(n)) %>%
# select(name, n, check)
# summary_logical_check %>% knit_big_table()
# }
```



### Other and translation
```{r}
if (check_for_others) {
  if (exists("questions")) {
    text_oth <- questions %>%
      filter(type == "text") %>%
      pull(name)
    
    other_log <- cleann %>%
        select(any_of(text_oth), !!sym(uuid_cleann)) %>%
        pivot_longer(cols = -!!sym(uuid_cleann)) %>%
        filter(!is.na(value)) %>%
        mutate(issue_type = "to could be recoded") %>%
        dplyr::rename(variable = name)
    
  } else {
    other_log <- cleaninginspectoR::find_other_responses(cleann) %>% 
      mutate(uuid = cleann[[uuid_cleann]][index]) 
  }
  list_log$other_log <- other_log
}
```
<br />
<br />
This is all the values from text questions.  

```{r}
if (check_for_others) {
  other_log %>%
    arrange(variable, value) %>%
    knit_big_table()
}
```
<br /> 
<br />
This is how many interviews per text question. 

```{r}
if (check_for_others) {
  other_log %>%
    group_by(variable) %>%
    tally(sort = T) %>%
    knit_big_table()
}
```
<br />
<br />
The values which are identicals. 

```{r}
if (check_for_others) {
  other_log %>%
    group_by(value) %>%
    tally(sort = T) %>%
    knit_big_table()
}
```


### Deletions
```{r}
number_cleann_in_dell <- "Check not performed"
number_dell_in_cleann <- "Check not performed"
number_difference_raw_clean_del <- "Check not performed"

if (check_for_deletions) {
  data.frame(n_raw = nrow(raww), 
             n_clean = nrow(cleann), 
             n_deleted = nrow(dell),
             sum_clean_del = nrow(cleann) + nrow(dell)
             )

  number_cleann_in_dell <- cleann[[uuid_cleann]] %in% dell[[uuid_dell]] %>% sum()
  number_dell_in_cleann <- dell[[uuid_dell]] %in% cleann[[uuid_cleann]] %>% sum()
  number_difference_raw_clean_del <-  abs(nrow(raww) - nrow(cleann) - nrow(dell))
  
  # if(number_cleann_in_dell != 0)
  #   values_cleann_in_dell <- 0
  #   values_cleann_in_dell <- setdiff(unique(cleann$uuidcleann), unique(dell$uuid_dell))
  #   values_cleann_in_dell <- data.frame(values_cleann_in_dell)
  # if(number_dell_in_cleann != 0)
  #   values_dell_in_cleann <- 0
  #   values_dell_in_cleann  <- setdiff(unique(dell$uuid_dell),unique(cleann$uuid_cleann))
  #   values_dell_in_cleann  <- data.frame(values_dell_in_cleann)
  # 
  # if(number_difference_raw_clean_del != 0)
  #   values_raww_in_clean <- 0
  #   values_raww_in_clean <- setdiff(unique(raww$uuid_raw),unique(cleann$uuid_cleann))
  #   values_raww_in_clean <- data.frame(values_raww_in_clean)
  #   values_raw_cleann_dell <- 0
  #   values_raw_cleann_dell <- setdiff(unique(values_raww_in_clean),unique(dell$uuid_dell))
  #   values_raw_cleann_dell <- data.frame(values_raw_cleann_dell)

}

    

```
Difference between raw, clean and deletion log:  `r number_difference_raw_clean_del`  
Number of uuid of clean in deleted : `r number_cleann_in_dell`  
Number of uuid of deleted in clean : `r number_dell_in_cleann`  
 
### Miscelleanous

## 2. Cleaning log {.tabset}

### Cleaning log

This section checks if the cleaning log has been applied.  

```{r, warning = F, echo = F}
cleaning_to_check <- data.frame()
number_cleaning_wrong <- "Check not performed"
number_cleaning_ambiugous <- "Check not performed"

if (check_for_cleaning_log) {
  old_new_values <- lapply(FUN = old_new,
                           X = split(logg, row.names(logg)), 
                           variable = var_logg,  
                           data_raw = raww,
                           data_clean = cleann, 
                           uuid_raw = uuid_raww,
                           uuid_clean = uuid_cleann, 
                           uuid_cleaning_log = uuid_logg 
                           ) %>%
    map(.f = function(xx){xx %>% mutate(across(.fns = as.character))}) %>%
    do.call(rbind, .)
  summary_cleaning <- logg %>% 
    left_join(old_new_values) %>% 
    mutate(check_raw = !!sym(old_value_logg) == value_raw,  
           check_clean = !!sym(new_value_logg) == value_clean)
  cleaning_to_check <- summary_cleaning %>% 
    filter((is.na(check_clean) & !!sym(new_value_logg) != "NA") | check_clean == F)  %>%
    select(all_of(c(uuid_logg, var_logg, old_value_logg, new_value_logg)), value_clean) %>%
    mutate(issue = "value from the cleaning log different from the value from the clean data.")
  
    number_cleaning_wrong <- summary_cleaning %>% filter(check_clean == F) %>% nrow()
    number_cleaning_ambiugous <- summary_cleaning %>% filter(is.na(check_clean), !!sym(new_value_logg) != "NA") %>% nrow()

}


```


```{r, warning = F}
if (check_for_cleaning_log) {
  cat("Number of entries in the cleaning log that have the same value in the clean data")
  summary_cleaning$check_clean %>% table(useNA = "ifany")
}
```
<br />
<br />
**How to read**  
- `r old_value_logg` refers to your cleaning log old value  
- `r new_value_logg` refers to your cleaning log new value  
- value_clean refers to the value in the clean dataset.
<br />
Cleaning log not applied for `r number_cleaning_wrong` entries.  
See below which ones.  

```{r, warning = F}
if (check_for_cleaning_log) {
  summary_cleaning %>% filter(check_clean == F) %>% 
    select(all_of(c(uuid_logg, var_logg, old_value_logg, new_value_logg)), value_clean) %>%
    knit_big_table()
}
```

<br />
<br />
Cleaning log ambigious for `r number_cleaning_ambiugous` entries.
You have a value in your cleaning log but it is NA in the dataset.  
This is a summary per question and per value. (feel free to ignore the obvious ones such as 999, deletion)
```{r, warning = F}
if (check_for_cleaning_log) {
  summary_cleaning %>% filter(is.na(check_clean), 
                               !!sym(new_value_logg) != "NA") %>% 
    group_by(!!sym(var_logg), !!sym(new_value_logg), value_clean) %>% tally(sort = T) %>% 
    knit_big_table()
}
```
<br />    
<br />
And the details here.  

```{r, warning = F}
if (check_for_cleaning_log) {
  cleaning_to_check_na <- summary_cleaning %>% filter(is.na(check_clean), 
                               !!sym(new_value_logg) != "NA") %>% 
    select(all_of(c(uuid_logg, var_logg, old_value_logg, new_value_logg)), value_clean)
  cleaning_to_check_na %>% 
    knit_big_table()
}
```


### Raw and Clean

This section checks if there are differences between the raw and the clean dataset that has not been reported.  

```{r, warning = F, echo = F, message = F}
raw_clean_to_check <- data.frame()
number_difference_raw_clean <- "Check not performed"


if (check_for_raw_clean_data) {
  raw_clean_compare <- compare_datasets(raww, uuid_raww, cleann, uuid_cleann) 

  raw_clean_compare$binding <- paste0(raw_clean_compare$uuid, "-/-", raw_clean_compare$question.name)
  
  raw_clean_compare$in_cleaning_log <- raw_clean_compare$binding %in% tolower(logg$binding)
}
```


```{r, warning = F, echo = F}
if (check_for_raw_clean_data) {
  number_diffrence_raw_clean <- sum(!raw_clean_compare$in_cleaning_log)
  cat("Number of differences between raw data and clean data that seems not be in the cleaning log")
  raw_clean_compare$in_cleaning_log %>% table(useNA = "ifany")
}
```
<br />
There are `r number_difference_raw_clean` changes not in the cleaning log.  
These have changed and does not appear in the cleaning log.  
```{r}
if (check_for_raw_clean_data) {
  raw_clean_to_check <- raw_clean_compare %>% 
    filter(comments == "change was made", in_cleaning_log == F) %>%
    mutate(issue = "value are different from the raw and clean and does not appear in the cleaning log") %>% 
    select(- comments)
  
  raw_clean_to_check %>% knit_big_table()
}
```

```{r}
library(writexl)
final_log <- list_log %>%
  map(.f = function(xx){xx %>% mutate(across(.fns = as.character))}) %>%
  do.call(bind_rows, .)

final_log %>% writexl::write_xlsx("outputs/review_cleaning_log_part1.xlsx")

cleaning_check <- list(cleaning_to_check = cleaning_to_check, 
                       raw_clean_to_check = raw_clean_to_check)

cleaning_check %>% writexl::write_xlsx("outputs/review_cleaning_part2.xlsx")

cleann %>%writexl::write_xlsx("outputs/dataset_with_check.xlsx")


```

